#T!!! = Task
#NM!!! = Not Match





---
title: "| Survival Analysis    \n| Data Scraping from Coldwell\n"
author: "April Leclair"
output:
  bookdown::tufte_html2:
    number_sections: no
    split_by: none
    toc: no
  bookdown::pdf_document2:
    latex_engine: pdflatex
    number_sections: no
    toc: no
  bookdown::tufte_handout2:
    latex_engine: xelatex
    number_sections: no
    toc: no
  bookdown::html_document2:
    number_sections: no
    split_by: none
    toc: no
---

```{r setup, include=FALSE}
library(tufte)
library(tidyverse)
library(devtools)
library(rvest)
library(stringr)
knitr::opts_chunk$set(tidy = FALSE, message=FALSE, cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)
```


## Retrieving Data Using `rvest` and the Selector Gadget

```{r cache=TRUE}
coldwell_tc1 <- read_html("https://www.coldwellbankerhomes.com/mn/saint-paul/recent-sales/kvc-17_1/uloc-75799_0/?sortId=2&offset=0&z=7&c=44.887712%2C-93.166878")
coldwell_tc2 <- read_html("https://www.coldwellbankerhomes.com/mn/minneapolis/recent-sales/kvc-17_1/uloc-76156_0/?sortId=2&offset=24")
coldwell_tc3 <- read_html("https://www.coldwellbankerhomes.com/mn/minneapolis/recent-sales/kvc-17_1/uloc-76156_0/?sortId=2&offset=48")
coldwell_tc4 <- read_html("https://www.coldwellbankerhomes.com/mn/minneapolis/recent-sales/kvc-17_1/uloc-76156_0/?sortId=2&offset=72")
coldwell_tc5 <- read_html("https://www.coldwellbankerhomes.com/mn/minneapolis/recent-sales/kvc-17_1/uloc-76156_0/?sortId=2&offset=96")
coldwell_tc6 <- read_html("https://www.coldwellbankerhomes.com/mn/minneapolis/recent-sales/kvc-17_1/uloc-76156_0/?sortId=2&offset=120")
coldwell_tc7 <- read_html("https://www.coldwellbankerhomes.com/mn/minneapolis/recent-sales/kvc-17_1/uloc-76156_0/?sortId=2&offset=144")
coldwell_tc8 <- read_html("https://www.coldwellbankerhomes.com/mn/minneapolis/recent-sales/kvc-17_1/uloc-76156_0/?sortId=2&offset=168")
coldwell_tc9 <- read_html("https://www.coldwellbankerhomes.com/mn/minneapolis/recent-sales/kvc-17_1/uloc-76156_0/?sortId=2&offset=192")
coldwell_tc10 <- read_html("https://www.coldwellbankerhomes.com/mn/minneapolis/recent-sales/kvc-17_1/uloc-76156_0/?sortId=2&offset=216")
coldwell_tc11 <- read_html("https://www.coldwellbankerhomes.com/mn/minneapolis/recent-sales/kvc-17_1/uloc-76156_0/?sortId=2&offset=240")
coldwell_tc12 <- read_html("https://www.coldwellbankerhomes.com/mn/minneapolis/recent-sales/kvc-17_1/uloc-76156_0/?sortId=2&offset=264")
coldwell_tc13 <- read_html("https://www.coldwellbankerhomes.com/mn/minneapolis/recent-sales/kvc-17_1/uloc-76156_0/?sortId=2&offset=288")
```

```{r}
coldwell_tc <- list(coldwell_tc1, coldwell_tc2, coldwell_tc3, coldwell_tc4, coldwell_tc5, coldwell_tc6, coldwell_tc7, coldwell_tc8, coldwell_tc9, coldwell_tc10, coldwell_tc11, coldwell_tc12, coldwell_tc13)
```


# Function: `onepg` scrapes a variable (determined by `path`) on a single page of a website
```{r}
onepg <- function(html, path) {
  varlst <- html %>%
      html_nodes(path) %>%
      html_text() 
  return(varlst) }
```


## Address
```{r}
address <- c(onepg(coldwell_tc1, ".street-address"), onepg(coldwell_tc2, ".street-address"), onepg(coldwell_tc3, ".street-address"), onepg(coldwell_tc4, ".street-address"), onepg(coldwell_tc5, ".street-address"), onepg(coldwell_tc6, ".street-address"), onepg(coldwell_tc7, ".street-address"), onepg(coldwell_tc8, ".street-address"), onepg(coldwell_tc9, ".street-address"), onepg(coldwell_tc10, ".street-address"), onepg(coldwell_tc11, ".street-address"), onepg(coldwell_tc12, ".street-address"), onepg(coldwell_tc13, ".street-address"))

address <- data.frame(address)
nrow(address)==24*13
```


## City & Zipcode
```{r}
cityzip <- c(onepg(coldwell_tc1, ".city-st-zip"), onepg(coldwell_tc2, ".city-st-zip"), onepg(coldwell_tc3, ".city-st-zip"), onepg(coldwell_tc4, ".city-st-zip"), onepg(coldwell_tc5, ".city-st-zip"), onepg(coldwell_tc6, ".city-st-zip"), onepg(coldwell_tc7, ".city-st-zip"), onepg(coldwell_tc8, ".city-st-zip"), onepg(coldwell_tc9, ".city-st-zip"), onepg(coldwell_tc10, ".city-st-zip"), onepg(coldwell_tc11, ".city-st-zip"), onepg(coldwell_tc12, ".city-st-zip"), onepg(coldwell_tc13, ".city-st-zip"))

cityzip <- data.frame(cityzip)
nrow(cityzip)==24*13

cityzip <- cityzip %>%
  mutate(city = as.factor(gsub(",.*$", "", cityzip)),
         zip = as.numeric(str_sub(cityzip, start= -5))) %>%
  select(city, zip)

# Note: 180 houses in Minnepolis & 132 houses in Saint Paul
```


# Number of Beds
```{r}
beds <- c(onepg(coldwell_tc1, ".beds .val"), onepg(coldwell_tc2, ".beds .val"), onepg(coldwell_tc3, ".beds .val"), onepg(coldwell_tc4, ".beds .val"), onepg(coldwell_tc5, ".beds .val"), onepg(coldwell_tc6, ".beds .val"), onepg(coldwell_tc7, ".beds .val"), onepg(coldwell_tc8, ".beds .val"), onepg(coldwell_tc9, ".beds .val"), onepg(coldwell_tc10, ".beds .val"), onepg(coldwell_tc11, ".beds .val"), onepg(coldwell_tc12, ".beds .val"), onepg(coldwell_tc13, ".beds .val"))

beds <- data.frame(beds)
nrow(beds)==24*13 #NM!!!
24*13-nrow(beds)


# coldwell_tc2 #23
# coldwell_tc3 #23
# coldwell_tc4 #23
# coldwell_tc6 #23
# coldwell_tc8 #22
# coldwell_tc12 #22
# coldwell_tc13 #23


beds2 <- (onepg(coldwell_tc2, ".beds .val"))
beds2[8:24] <- c(1, beds2[8:23])

beds3 <- (onepg(coldwell_tc3, ".beds .val"))
beds3[22:24] <- c(1, beds2[22:23])

beds4 <- (onepg(coldwell_tc4, ".beds .val"))
beds4[19:24] <- c(1, beds4[22:23])

beds6 <- (onepg(coldwell_tc6, ".beds .val"))
beds6[17:24] <- c(1, beds6[17:23])

beds8 <- (onepg(coldwell_tc8, ".beds .val"))
beds8[3:23] <- c(1, beds8[3:22])
beds8[15:24] <- c(1, beds8[15:23])

beds12 <- (onepg(coldwell_tc12, ".beds .val"))
beds12[5:23] <- c(1, beds12[5:22])
beds12[10:24] <- c(1, beds12[10:23])

beds13 <- (onepg(coldwell_tc13, ".beds .val"))
beds13[12:24] <- c(1, beds13[12:23])

beds <- c(onepg(coldwell_tc1, ".beds .val"), beds2, beds3, beds4, onepg(coldwell_tc5, ".beds .val"), beds6, onepg(coldwell_tc7, ".beds .val"), beds8, onepg(coldwell_tc9, ".beds .val"), onepg(coldwell_tc10, ".beds .val"), onepg(coldwell_tc11, ".beds .val"), beds12, beds13)

beds <- data.frame(beds)
nrow(beds)==24*13

test <- data.frame(cbind(address, beds))
test %>% filter(beds==1)

rm(test, beds2, beds3, beds4, beds6, beds8, beds12, beds13)
```


# Number of Full Bath
```{r}
bathf <- c(onepg(coldwell_tc1, "li:nth-child(2) .val"), onepg(coldwell_tc2, "li:nth-child(2) .val"), onepg(coldwell_tc3, "li:nth-child(2) .val"), onepg(coldwell_tc4, "li:nth-child(2) .val"), onepg(coldwell_tc5, "li:nth-child(2) .val"), onepg(coldwell_tc6, "li:nth-child(2) .val"), onepg(coldwell_tc7, "li:nth-child(2) .val"), onepg(coldwell_tc8, "li:nth-child(2) .val"), onepg(coldwell_tc9, "li:nth-child(2) .val"), onepg(coldwell_tc10, "li:nth-child(2) .val"), onepg(coldwell_tc11, "li:nth-child(2) .val"), onepg(coldwell_tc12, "li:nth-child(2) .val"), onepg(coldwell_tc13, "li:nth-child(2) .val"))

bathf <- data.frame(bathf)
nrow(bathf)==24*13
```


# Number of Car carg
```{r}
carg <- c(onepg(coldwell_tc1, ".car-garage .val"), onepg(coldwell_tc2, ".car-garage .val"), onepg(coldwell_tc3, ".car-garage .val"), onepg(coldwell_tc4, ".car-garage .val"), onepg(coldwell_tc5, ".car-garage .val"), onepg(coldwell_tc6, ".car-garage .val"), onepg(coldwell_tc7, ".car-garage .val"), onepg(coldwell_tc8, ".car-garage .val"), onepg(coldwell_tc9, ".car-garage .val"), onepg(coldwell_tc10, ".car-garage .val"), onepg(coldwell_tc11, ".car-garage .val"), onepg(coldwell_tc12, ".car-garage .val"), onepg(coldwell_tc13, ".car-garage .val"))


carg <- data.frame(carg)
nrow(carg)==24*13 #NM!!!
24*13-nrow(carg)  


carg2 <- (onepg(coldwell_tc2, ".car-carg.val"))
carg2[8:23] <- c(0, carg2[8:22])
carg2[9:24] <- c(0, carg2[9:23])

carg3 <- (onepg(coldwell_tc3, ".car-carg.val"))
carg3[6:24] <- c(0, carg3[6:23])

carg4 <- (onepg(coldwell_tc4, ".car-carg.val"))
carg4[2:23] <- c(0, carg4[2:22])
carg4[21:24] <- c(0, carg4[21:23])

carg5 <- (onepg(coldwell_tc5, ".car-carg.val"))
carg5[14:23] <- c(0, carg5[14:22])
carg5[21:24] <- c(0, carg5[21:23])

carg6 <- (onepg(coldwell_tc6, ".car-carg.val"))
carg6[6:23] <- c(0, carg6[6:22])
carg6[10:24] <- c(0, carg6[10:23])

carg7 <- (onepg(coldwell_tc7, ".car-carg.val"))
carg7[2:19] <- c(0, carg7[2:18])
carg7[4:20] <- c(0, carg7[4:19])
carg7[5:21] <- c(0, carg7[5:20])
carg7[7:22] <- c(0, carg7[7:21])
carg7[8:23] <- c(0, carg7[8:22])
carg7[12:24] <- c(0, carg7[12:23])

carg8 <- (onepg(coldwell_tc8, ".car-carg.val"))
carg8[16:22] <- c(0, carg8[16:21])
carg8[20:23] <- c(0, carg8[20:22])
carg8[23:24] <- c(0, carg8[23:23])

carg9 <- (onepg(coldwell_tc9, ".car-carg.val"))
carg9[8:24] <- c(0, carg9[8:23])

carg10 <- (onepg(coldwell_tc10, ".car-carg.val"))
carg10[2:19] <- c(0, carg10[2:18])
carg10[8:20] <- c(0, carg10[8:19])
carg10[11:21] <- c(0, carg10[11:20])
carg10[14:22] <- c(0, carg10[14:21])
carg10[16:23] <- c(0, carg10[16:22])
carg10[18:24] <- c(0, carg10[18:23])

carg12 <- (onepg(coldwell_tc12, ".car-carg.val"))
carg12[6:22] <- c(0, carg12[6:21])
carg12[19:23] <- c(0, carg12[19:22])
carg12[20:24] <- c(0, carg12[20:23])


carg <- c(onepg(coldwell_tc1, ".car-carg .val"), carg2, carg3, carg4, carg5, carg6, carg7, carg8, carg9, carg10, onepg(coldwell_tc11, ".car-carg .val"), carg12, onepg(coldwell_tc13, ".car-carg .val"))

carg <- data.frame(carg)
nrow(carg)==24*13

test <- data.frame(cbind(address, carg))
test %>% filter(carg==0)
```


# Price
```{r}
price <- c(onepg(coldwell_tc1, ".price-normal"), onepg(coldwell_tc2, ".price-normal"), onepg(coldwell_tc3, ".price-normal"), onepg(coldwell_tc4, ".price-normal"), onepg(coldwell_tc5, ".price-normal"), onepg(coldwell_tc6, ".price-normal"), onepg(coldwell_tc7, ".price-normal"), onepg(coldwell_tc8, ".price-normal"), onepg(coldwell_tc9, ".price-normal"), onepg(coldwell_tc10, ".price-normal"), onepg(coldwell_tc11, ".price-normal"), onepg(coldwell_tc12, ".price-normal"), onepg(coldwell_tc13, ".price-normal"))

price <- data.frame(price)
price <- price %>%
  mutate(price = as.numeric(gsub("\\$|,", "", as.character(price))))

nrow(price)==24*13
```


# Make the Data Table
```{r}
house <- cbind(address, cityzip, beds, bathf, price)
```


## Bind multiple pages
```{r}
# Make a variable `pid <- "pid_21803186/"` this pair #T!!!
# Make a variable `lid <- "minneapolis/4540-penn-avenue-n/" + pid` for each address #T!!!
link <- "https://www.coldwellbankerhomes.com/mn/minneapolis/4540-penn-avenue-n/pid_21803186/"
link2 <- "https://www.coldwellbankerhomes.com/mn/saint-paul/766-jackson-street/pid_20380688/"
coldwell_ind <- read_html(link)


pid <- "pid_21803186/"
# FOR ONE PID FOR NOW #T!!!
# Makes a link id for each house (a part of the full link)
lstlid <- house %>%
  mutate(first = paste(tolower(gsub("\\s", "-", city)), "/"),
         second = paste(tolower(gsub("\\s", "-", address)), "/"),
         lid = gsub("\\s", "", paste(first, second, pid))) %>%
  select(-first, -second)



# Makes a list of web link for each house
lstlink <- function(lid) {
  link <- "https://www.coldwellbankerhomes.com/mn/"
  str <- ""
  lst <- lst()
  for (i in 1: length(lid)) {
    link <- gsub("\\s", "", paste(link, lid[i]))
    lst[i] <- link
  }
  return(lst)
}


# Personal note: test for adding strings
t <- lst()
d <- "dd" 
t[1] <- gsub("\\s", "", paste("rg", d))
```


pid_21201050


```{r}
address
pid <- c("pid_21201050", "pid_21134768", "pid3")

cbind(address,pid)
```

# Paths saved
```{r}
soldat <- "ul:nth-child(4) li:nth-child(2)"
listedat <- "ul:nth-child(4) li:nth-child(3)"
schname <- ".school-name"
schaddr <- ".school-address"
schgd <- ".property-detail-school-listing .school-data-grades"
schdist <- ".property-detail-school-listing .school-data-distance"
schdstr <- ".split-2 .col:nth-child(2) li" #?
```



```{r}
coldwell_ind <- read_html(link)
onepg(coldwell_ind, soldat)
```


