#T!!! = Task
#NM!!! = Not Match





---
title: "| Survival Analysis    \n| Data Scraping from Coldwell\n"
author: "April Leclair"
output:
  bookdown::tufte_html2:
    number_sections: no
    split_by: none
    toc: no
  bookdown::pdf_document2:
    latex_engine: pdflatex
    number_sections: no
    toc: no
  bookdown::tufte_handout2:
    latex_engine: xelatex
    number_sections: no
    toc: no
  bookdown::html_document2:
    number_sections: no
    split_by: none
    toc: no
---

```{r setup, include=FALSE, message=FALSE}
library(tufte)
library(tidyverse)
library(devtools)
library(rvest)
library(stringr)
knitr::opts_chunk$set(tidy = FALSE, message=FALSE, cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)
```


# Read the Saved HTML
```{r}
# file.choose()
coldwell_tc1a <- "../Source/Minneapolis, MN Recent Home Sales (Page 1).html"
coldwell_tc2a <- "../Source/Minneapolis, MN Recent Home Sales (Page 2).html"
coldwell_tc3a <- "../Source/Minneapolis, MN Recent Home Sales (Page 3).html"
coldwell_tc4a <- "../Source/Minneapolis, MN Recent Home Sales (Page 4).html"
coldwell_tc5a <- "../Source/Minneapolis, MN Recent Home Sales (Page 5).html"
coldwell_tc6a <- "../Source/Minneapolis, MN Recent Home Sales (Page 6).html"
coldwell_tc7a <- "../Source/Minneapolis, MN Recent Home Sales (Page 7).html"
coldwell_tc8a <- "../Source/Minneapolis, MN Recent Home Sales (Page 8).html"
coldwell_tc9a <- "../Source/Minneapolis, MN Recent Home Sales (Page 9).html"
coldwell_tc10a <- "../Source/Minneapolis, MN Recent Home Sales (Page 10).html"
coldwell_tc11a <- "../Source/Minneapolis, MN Recent Home Sales (Page 11).html"
coldwell_tc12a <- "../Source/Minneapolis, MN Recent Home Sales (Page 12).html"
coldwell_tc13a <- "../Source/Minneapolis, MN Recent Home Sales (Page 13).html"
```

```{r}
coldwell_tc1 <- read_html(coldwell_tc1a)
coldwell_tc2 <- read_html(coldwell_tc2a)
coldwell_tc3 <- read_html(coldwell_tc3a)
coldwell_tc4 <- read_html(coldwell_tc4a)
coldwell_tc5 <- read_html(coldwell_tc5a)
coldwell_tc6 <- read_html(coldwell_tc6a)
coldwell_tc7 <- read_html(coldwell_tc7a) 
coldwell_tc8 <- read_html(coldwell_tc8a)
coldwell_tc9 <- read_html(coldwell_tc9a)
coldwell_tc10 <- read_html(coldwell_tc10a)
coldwell_tc11 <- read_html(coldwell_tc11a)
coldwell_tc12 <- read_html(coldwell_tc12a)
coldwell_tc13 <- read_html(coldwell_tc13a)
```


# Create a Scraping Function 
```{r}
# `scrape` scrapes a variable (determined by `path`) on a single page of a website
scrape <- function(readhtml, path) {
  varlst <- readhtml %>%
    html_nodes(path) %>%
    html_text() 
  return(varlst) }
```



# Variables

## Address
```{r}
address <- c(scrape(coldwell_tc1, ".street-address"), scrape(coldwell_tc2, ".street-address"), scrape(coldwell_tc3, ".street-address"), scrape(coldwell_tc4, ".street-address"), scrape(coldwell_tc5, ".street-address"), scrape(coldwell_tc6, ".street-address"), scrape(coldwell_tc7, ".street-address"), scrape(coldwell_tc8, ".street-address"), scrape(coldwell_tc9, ".street-address"), scrape(coldwell_tc10, ".street-address"), scrape(coldwell_tc11, ".street-address"), scrape(coldwell_tc12, ".street-address"), scrape(coldwell_tc13, ".street-address"))

address <- data.frame(address) 
address <- address %>% mutate(address = str_trim(address))
nrow(address)==24*13
```


## City & Zipcode
```{r}
cityzip <- c(scrape(coldwell_tc1, ".city-st-zip"), scrape(coldwell_tc2, ".city-st-zip"), scrape(coldwell_tc3, ".city-st-zip"), scrape(coldwell_tc4, ".city-st-zip"), scrape(coldwell_tc5, ".city-st-zip"), scrape(coldwell_tc6, ".city-st-zip"), scrape(coldwell_tc7, ".city-st-zip"), scrape(coldwell_tc8, ".city-st-zip"), scrape(coldwell_tc9, ".city-st-zip"), scrape(coldwell_tc10, ".city-st-zip"), scrape(coldwell_tc11, ".city-st-zip"), scrape(coldwell_tc12, ".city-st-zip"), scrape(coldwell_tc13, ".city-st-zip"))

cityzip <- data.frame(cityzip)
nrow(cityzip)==24*13

cityzip <- cityzip %>%
  mutate(city = as.factor(gsub(",.*$", "", cityzip)),
         zip = as.numeric(str_sub(cityzip, start= -5))) %>%
  select(city, zip)

# Note: 171 houses in Minnepolis & 141 houses in Saint Paul
# sum(cityzip$city=="Minneapolis")
# sum(cityzip$city=="Saint Paul")
```


## Number of Beds
```{r}
beds <- c(scrape(coldwell_tc1, ".beds .val"), scrape(coldwell_tc2, ".beds .val"), scrape(coldwell_tc3, ".beds .val"), scrape(coldwell_tc4, ".beds .val"), scrape(coldwell_tc5, ".beds .val"), scrape(coldwell_tc6, ".beds .val"), scrape(coldwell_tc7, ".beds .val"), scrape(coldwell_tc8, ".beds .val"), scrape(coldwell_tc9, ".beds .val"), scrape(coldwell_tc10, ".beds .val"), scrape(coldwell_tc11, ".beds .val"), scrape(coldwell_tc12, ".beds .val"), scrape(coldwell_tc13, ".beds .val"))

beds <- data.frame(beds)
nrow(beds)==24*13 
24*13-nrow(beds) #8

# length(scrape(coldwell_tc2, ".beds .val"))
# scrape(coldwell_tc2, ".street-address")

beds2 <- (scrape(coldwell_tc2, ".beds .val"))
beds2[20:24] <- c(1, beds2[20:23]) #3415

beds4 <- (scrape(coldwell_tc4, ".beds .val"))
beds4[11:24] <- c(1, beds4[11:23])

beds5 <- (scrape(coldwell_tc5, ".beds .val"))
beds5[8:24] <- c(1, beds5[8:23])

beds7 <- (scrape(coldwell_tc7, ".beds .val"))
beds7[6:24] <- c(1, beds7[6:23])

beds8 <- (scrape(coldwell_tc8, ".beds .val"))
beds8[16:24] <- c(1, beds8[16:23])

beds9 <- (scrape(coldwell_tc9, ".beds .val"))
beds9[4:24] <- c(1, beds9[4:23])

beds12 <- (scrape(coldwell_tc12, ".beds .val"))
beds12[18:23] <- c(1, beds12[18:22])
beds12[23:24] <- c(1, beds12[23:23])

beds <- c(scrape(coldwell_tc1, ".beds .val"), beds2, scrape(coldwell_tc3, ".beds .val"), beds4, beds5, scrape(coldwell_tc6, ".beds .val"), beds7, beds8, beds9, scrape(coldwell_tc10, ".beds .val"), scrape(coldwell_tc11, ".beds .val"), beds12, scrape(coldwell_tc13, ".beds .val"))

beds <- data.frame(beds)
beds <- beds %>% mutate(beds = as.numeric(beds))
nrow(beds)==24*13

# test <- data.frame(cbind(address, beds))
# test %>% filter(beds==1)
# rm(test)
rm(beds2, beds4, beds5, beds7, beds8, beds9, beds12)
```


## Number of Full Bath
```{r}
bathf <- c(scrape(coldwell_tc1, "li:nth-child(2) .val"), scrape(coldwell_tc2, "li:nth-child(2) .val"), scrape(coldwell_tc3, "li:nth-child(2) .val"), scrape(coldwell_tc4, "li:nth-child(2) .val"), scrape(coldwell_tc5, "li:nth-child(2) .val"), scrape(coldwell_tc6, "li:nth-child(2) .val"), scrape(coldwell_tc7, "li:nth-child(2) .val"), scrape(coldwell_tc8, "li:nth-child(2) .val"), scrape(coldwell_tc9, "li:nth-child(2) .val"), scrape(coldwell_tc10, "li:nth-child(2) .val"), scrape(coldwell_tc11, "li:nth-child(2) .val"), scrape(coldwell_tc12, "li:nth-child(2) .val"), scrape(coldwell_tc13, "li:nth-child(2) .val"))

bathf <- data.frame(bathf)
bathf <- bathf %>% mutate(bathf = as.numeric(bathf))
nrow(bathf)==24*13
```


## Number of Partial Bath
```{r}
bathp1 <- c(1, 2, 1, 1, 1, 0, 1, 1, 1, 0, 0, 1, 1, 0, 0, 0, 2, 1, 1, 2, 0, 1, 1, 0)
bathp2 <- c(0, 1, 0, 1, 1, 1, 0, 1, 0, 2, 0, 1, 0, 0, 1, 1, 1, 1, 1, 0, 1, 1, 1, 1)
bathp3 <- c(0, 1, 1, 3, 0, 1, 0, 1, 0, 2, 1, 0, 1, 0, 0, 1, 1, 1, 0, 1, 1, 0, 2, 0)
bathp4 <- c(2, 1, 0, 1, 2, 0, 2, 0, 1, 1, 0, 1, 0, 0, 0, 0, 0, 1, 0, 1, 1, 1, 1, 0)
bathp5 <- c(0, 3, 0, 1, 1, 0, 1, 0, 1, 2, 2, 1, 0, 0, 1, 1, 2, 3, 1, 1, 1, 1, 1, 1)
bathp6 <- c(2, 1, 0, 0, 0, 0, 1, 1, 1, 0, 0, 0, 0, 0, 1, 2, 0, 0, 1, 0, 1, 0, 0, 0)
bathp7 <- c(2, 1, 0, 0, 1, 0, 1, 1, 1, 1, 1, 0, 1, 1, 0, 0, 0, 1, 1, 0, 0, 1, 2, 1)
bathp8 <- c(1, 0, 0, 0, 0, 0, 2, 0, 0, 0, 0, 0, 3, 1, 0, 0, 0, 1, 0, 0, 0, 1, 2, 1)
bathp9 <- c(1, 0, 3, 0, 0, 0, 1, 2, 0, 1, 1, 0, 0, 1, 0, 0, 0, 1, 1, 0, 0, 0, 1, 0) 
bathp10 <- c(1, 0, 1, 0, 1, 1, 1, 1, 0, 1, 0, 1, 0, 0, 0, 1, 0, 1, 1, 1, 0, 1, 0, 0)
bathp11 <- c(2, 1, 0, 1, 5, 1, 1, 1, 1, 2, 0, 0, 1, 0, 1, 1, 1, 0, 0, 1, 2, 1, 0, 1)
bathp12 <- c(2, 2, 0, 2, 1, 1, 1, 1, 0, 0, 0, 1, 0, 0, 1, 1, 1, 0, 0, 0, 2, 0, 2, 0)
bathp13 <- c(0, 1, 3, 0, 1, 1, 0, 0, 0, 0, 0, 1, 0, 0, 2, 2, 1, 0, 0, 1, 1, 0, 1, 3)

# test <- data.frame(cbind(address, bathp))
# test
# rm(test)

bathp <- c(bathp1, bathp2, bathp3, bathp4, bathp5, bathp6, bathp7, bathp8, bathp9, bathp10, bathp11, bathp12, bathp13)
bathp <- data.frame(bathp)
nrow(bathp)==24*13
rm(bathp1, bathp2, bathp3, bathp4, bathp5, bathp6, bathp7, bathp8, bathp9, bathp10, bathp11, bathp12, bathp13)
```


## Number of Car Garage
```{r}
carg <- c(scrape(coldwell_tc1, ".car-garage .val"), scrape(coldwell_tc2, ".car-garage .val"), scrape(coldwell_tc3, ".car-garage .val"), scrape(coldwell_tc4, ".car-garage .val"), scrape(coldwell_tc5, ".car-garage .val"), scrape(coldwell_tc6, ".car-garage .val"), scrape(coldwell_tc7, ".car-garage .val"), scrape(coldwell_tc8, ".car-garage .val"), scrape(coldwell_tc9, ".car-garage .val"), scrape(coldwell_tc10, ".car-garage .val"), scrape(coldwell_tc11, ".car-garage .val"), scrape(coldwell_tc12, ".car-garage .val"), scrape(coldwell_tc13, ".car-garage .val"))

carg <- data.frame(carg)
nrow(carg)==24*13 
24*13-nrow(carg) # 29  

# length(scrape(coldwell_tc1, ".car-garage .val"))
# scrape(coldwell_tc6, ".street-address")

carg1 <- (scrape(coldwell_tc1, ".car-garage .val"))
carg1[18:24] <- c(0, carg1[18:23])

carg2 <- (scrape(coldwell_tc2, ".car-garage .val"))
carg2[20:23] <- c(0, carg2[20:22])
carg2[21:24] <- c(0, carg2[21:23])

carg3 <- (scrape(coldwell_tc3, ".car-garage .val"))
carg3[19:24] <- c(0, carg3[19:23])

carg4 <- (scrape(coldwell_tc4, ".car-garage .val"))
carg4[15:24] <- c(0, carg4[15:23])

carg5 <- (scrape(coldwell_tc5, ".car-garage .val"))
carg5[10:24] <- c(0, carg5[10:23]) 

carg6 <- (scrape(coldwell_tc6, ".car-garage .val"))
carg6[3:21] <- c(0, carg6[3:20]) 
carg6[10:22] <- c(0, carg6[10:21]) 
carg6[19:23] <- c(0, carg6[19:22]) 
carg6[23:24] <- c(0, carg6[23:23]) 

carg7 <- (scrape(coldwell_tc7, ".car-garage .val")) 
carg7[15:20] <- c(0, carg7[15:19]) 
carg7[17:21] <- c(0, carg7[17:20]) 
carg7[18:22] <- c(0, carg7[18:21]) 
carg7[20:23] <- c(0, carg7[20:22]) 
carg7[21:24] <- c(0, carg7[21:23]) 

carg8 <- (scrape(coldwell_tc8, ".car-garage .val")) 
carg8[1:24] <- c(0, carg8[1:23]) 

carg9 <- (scrape(coldwell_tc9, ".car-garage .val"))
carg9[5:21] <- c(0, carg9[5:20]) 
carg9[9:22] <- c(0, carg9[9:21]) 
carg9[12:23] <- c(0, carg9[12:22]) 
carg9[21:24] <- c(0, carg9[21:23]) 

carg10 <- (scrape(coldwell_tc10, ".car-garage .val"))
carg10[15:22] <- c(0, carg10[15:21])
carg10[21:23] <- c(0, carg10[21:22])
carg10[24:24] <- c(0)

carg11 <- (scrape(coldwell_tc11, ".car-garage .val"))
carg11[3:22] <- c(0, carg11[3:21])
carg11[5:23] <- c(0, carg11[5:22])
carg11[7:24] <- c(0, carg11[7:23])

carg12 <- (scrape(coldwell_tc12, ".car-garage .val"))
carg12[19:24] <- c(0, carg12[19:23])

carg13 <- (scrape(coldwell_tc13, ".car-garage .val"))
carg13[8:23] <- c(0, carg13[8:22])
carg13[9:24] <- c(0, carg13[9:23])

carg <- c(carg1, carg2, carg3, carg4, carg5, carg6, carg7, carg8, carg9, carg10, carg11, carg12, carg13)

carg <- data.frame(carg)
carg <- carg %>% mutate(carg = as.numeric(carg))
nrow(carg)==24*13

# test <- data.frame(cbind(address, carg))
# test %>% filter(carg==0)
# rm(test)

rm(carg1, carg2, carg3, carg4, carg5, carg6, carg7, carg8, carg9, carg10, carg11, carg12, carg13)
```


# Price
```{r}
price <- c(scrape(coldwell_tc1, ".price-normal"), scrape(coldwell_tc2, ".price-normal"), scrape(coldwell_tc3, ".price-normal"), scrape(coldwell_tc4, ".price-normal"), scrape(coldwell_tc5, ".price-normal"), scrape(coldwell_tc6, ".price-normal"), scrape(coldwell_tc7, ".price-normal"), scrape(coldwell_tc8, ".price-normal"), scrape(coldwell_tc9, ".price-normal"), scrape(coldwell_tc10, ".price-normal"), scrape(coldwell_tc11, ".price-normal"), scrape(coldwell_tc12, ".price-normal"), scrape(coldwell_tc13, ".price-normal"))

price <- data.frame(price)
price <- price %>%
  mutate(price = as.numeric(gsub("\\$|,", "", as.character(price))))
nrow(price)==24*13
```


# Make the Data Table
```{r}
house <- cbind(address, cityzip, beds, bathf, bathp, carg, price)
save(house, file = "../Source/house.csv")
```


############################## BELOW IS INCOMPLETE #############################################

## Bind multiple pages
```{r}
# Make a variable `pid <- "pid_21803186/"` this pair #T!!!
# Make a variable `lid <- "minneapolis/4540-penn-avenue-n/" + pid` for each address #T!!!
link <- "https://www.coldwellbankerhomes.com/mn/minneapolis/4540-penn-avenue-n/pid_21803186/"
link2 <- "https://www.coldwellbankerhomes.com/mn/saint-paul/766-jackson-street/pid_20380688/"
coldwell_ind <- read_html(link)


pid <- "pid_21803186/"
# FOR ONE PID FOR NOW #T!!!
# Makes a link id for each house (a part of the full link)
lstlid <- house %>%
  mutate(first = paste(tolower(gsub("\\s", "-", city)), "/"),
         second = paste(tolower(gsub("\\s", "-", address)), "/"),
         lid = gsub("\\s", "", paste(first, second, pid))) %>%
  select(-first, -second)



# Makes a list of web link for each house
lstlink <- function(lid) {
  link <- "https://www.coldwellbankerhomes.com/mn/"
  str <- ""
  lst <- lst()
  for (i in 1: length(lid)) {
    link <- gsub("\\s", "", paste(link, lid[i]))
    lst[i] <- link
  }
  return(lst)
}


# Personal note: test for adding strings
t <- lst()
d <- "dd" 
t[1] <- gsub("\\s", "", paste("rg", d))
```

```{r}
address
pid <- c("pid_21201050", "pid_21134768", "pid3")

cbind(address,pid)
```

# Paths saved
```{r}
soldat <- "ul:nth-child(4) li:nth-child(2)"
listedat <- "ul:nth-child(4) li:nth-child(3)"
schname <- ".school-name"
schaddr <- ".school-address"
schgd <- ".property-detail-school-listing .school-data-grades"
schdist <- ".property-detail-school-listing .school-data-distance"
schdstr <- ".split-2 .col:nth-child(2) li" #?
```



```{r}
coldwell_ind <- read_html(link)
scrape(coldwell_ind, soldat)
```


