---
title: "| Survival Analysis    \n| Data Scraping from Coldwell\n"
author: "April Leclair"
output:
  bookdown::tufte_html2:
    number_sections: no
    split_by: none
    toc: no
  bookdown::pdf_document2:
    latex_engine: pdflatex
    number_sections: no
    toc: no
  bookdown::tufte_handout2:
    latex_engine: xelatex
    number_sections: no
    toc: no
  bookdown::html_document2:
    number_sections: no
    split_by: none
    toc: no
---
 
```{r setup, include=FALSE, message=FALSE}
library(tufte)
library(tidyverse)
library(devtools)
library(rvest)
library(stringr)
knitr::opts_chunk$set(tidy = FALSE, message=FALSE, cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)
```


# Read the Saved HTML
```{r}
# file.choose()
coldwell_tc1a <- "../Source/Minneapolis, MN Recent Home Sales (Page 1).html"
coldwell_tc2a <- "../Source/Minneapolis, MN Recent Home Sales (Page 2).html"
coldwell_tc3a <- "../Source/Minneapolis, MN Recent Home Sales (Page 3).html"
coldwell_tc4a <- "../Source/Minneapolis, MN Recent Home Sales (Page 4).html"
coldwell_tc5a <- "../Source/Minneapolis, MN Recent Home Sales (Page 5).html"
coldwell_tc6a <- "../Source/Minneapolis, MN Recent Home Sales (Page 6).html"
coldwell_tc7a <- "../Source/Minneapolis, MN Recent Home Sales (Page 7).html"
coldwell_tc8a <- "../Source/Minneapolis, MN Recent Home Sales (Page 8).html"
coldwell_tc9a <- "../Source/Minneapolis, MN Recent Home Sales (Page 9).html"
coldwell_tc10a <- "../Source/Minneapolis, MN Recent Home Sales (Page 10).html"
coldwell_tc11a <- "../Source/Minneapolis, MN Recent Home Sales (Page 11).html"
coldwell_tc12a <- "../Source/Minneapolis, MN Recent Home Sales (Page 12).html"
coldwell_tc13a <- "../Source/Minneapolis, MN Recent Home Sales (Page 13).html"
```

```{r}
coldwell_tc1 <- read_html(coldwell_tc1a)
coldwell_tc2 <- read_html(coldwell_tc2a)
coldwell_tc3 <- read_html(coldwell_tc3a)
coldwell_tc4 <- read_html(coldwell_tc4a)
coldwell_tc5 <- read_html(coldwell_tc5a)
coldwell_tc6 <- read_html(coldwell_tc6a)
coldwell_tc7 <- read_html(coldwell_tc7a) 
coldwell_tc8 <- read_html(coldwell_tc8a)
coldwell_tc9 <- read_html(coldwell_tc9a)
coldwell_tc10 <- read_html(coldwell_tc10a)
coldwell_tc11 <- read_html(coldwell_tc11a)
coldwell_tc12 <- read_html(coldwell_tc12a)
coldwell_tc13 <- read_html(coldwell_tc13a)
```


# Function: Create a Scraping Function 
```{r}
# `scrape` scrapes a variable (determined by `path`) on a single page of a website
scrape <- function(readhtml, path) {
  varlst <- readhtml %>%
    html_nodes(path) %>%
    html_text() 
  return(varlst) }
```



# Variables

## Address
```{r}
address <- c(scrape(coldwell_tc1, ".street-address"), scrape(coldwell_tc2, ".street-address"), scrape(coldwell_tc3, ".street-address"), scrape(coldwell_tc4, ".street-address"), scrape(coldwell_tc5, ".street-address"), scrape(coldwell_tc6, ".street-address"), scrape(coldwell_tc7, ".street-address"), scrape(coldwell_tc8, ".street-address"), scrape(coldwell_tc9, ".street-address"), scrape(coldwell_tc10, ".street-address"), scrape(coldwell_tc11, ".street-address"), scrape(coldwell_tc12, ".street-address"), scrape(coldwell_tc13, ".street-address"))

address <- data.frame(address) 
address <- address %>% mutate(address = str_trim(address))
nrow(address)==24*13
```

## City & Zipcode
```{r}
cityzip <- c(scrape(coldwell_tc1, ".city-st-zip"), scrape(coldwell_tc2, ".city-st-zip"), scrape(coldwell_tc3, ".city-st-zip"), scrape(coldwell_tc4, ".city-st-zip"), scrape(coldwell_tc5, ".city-st-zip"), scrape(coldwell_tc6, ".city-st-zip"), scrape(coldwell_tc7, ".city-st-zip"), scrape(coldwell_tc8, ".city-st-zip"), scrape(coldwell_tc9, ".city-st-zip"), scrape(coldwell_tc10, ".city-st-zip"), scrape(coldwell_tc11, ".city-st-zip"), scrape(coldwell_tc12, ".city-st-zip"), scrape(coldwell_tc13, ".city-st-zip"))

cityzip <- data.frame(cityzip)
nrow(cityzip)==24*13

cityzip <- cityzip %>%
  mutate(city = as.factor(gsub(",.*$", "", cityzip)),
         zip = as.numeric(str_sub(cityzip, start= -5))) %>%
  select(city, zip)

# Note: 171 houses in Minnepolis & 141 houses in Saint Paul
# sum(cityzip$city=="Minneapolis")
# sum(cityzip$city=="Saint Paul")
```

## Number of Beds
```{r}
beds <- c(scrape(coldwell_tc1, ".beds .val"), scrape(coldwell_tc2, ".beds .val"), scrape(coldwell_tc3, ".beds .val"), scrape(coldwell_tc4, ".beds .val"), scrape(coldwell_tc5, ".beds .val"), scrape(coldwell_tc6, ".beds .val"), scrape(coldwell_tc7, ".beds .val"), scrape(coldwell_tc8, ".beds .val"), scrape(coldwell_tc9, ".beds .val"), scrape(coldwell_tc10, ".beds .val"), scrape(coldwell_tc11, ".beds .val"), scrape(coldwell_tc12, ".beds .val"), scrape(coldwell_tc13, ".beds .val"))

beds <- data.frame(beds)
nrow(beds)==24*13 
24*13-nrow(beds) #8

# length(scrape(coldwell_tc2, ".beds .val"))
# scrape(coldwell_tc2, ".street-address")

beds2 <- (scrape(coldwell_tc2, ".beds .val"))
beds2[20:24] <- c(1, beds2[20:23]) #3415

beds4 <- (scrape(coldwell_tc4, ".beds .val"))
beds4[11:24] <- c(1, beds4[11:23])

beds5 <- (scrape(coldwell_tc5, ".beds .val"))
beds5[8:24] <- c(1, beds5[8:23])

beds7 <- (scrape(coldwell_tc7, ".beds .val"))
beds7[6:24] <- c(1, beds7[6:23])

beds8 <- (scrape(coldwell_tc8, ".beds .val"))
beds8[16:24] <- c(1, beds8[16:23])

beds9 <- (scrape(coldwell_tc9, ".beds .val"))
beds9[4:24] <- c(1, beds9[4:23])

beds12 <- (scrape(coldwell_tc12, ".beds .val"))
beds12[18:23] <- c(1, beds12[18:22])
beds12[23:24] <- c(1, beds12[23:23])

beds <- c(scrape(coldwell_tc1, ".beds .val"), beds2, scrape(coldwell_tc3, ".beds .val"), beds4, beds5, scrape(coldwell_tc6, ".beds .val"), beds7, beds8, beds9, scrape(coldwell_tc10, ".beds .val"), scrape(coldwell_tc11, ".beds .val"), beds12, scrape(coldwell_tc13, ".beds .val"))

beds <- data.frame(beds)
beds <- beds %>% mutate(beds = as.numeric(beds))
nrow(beds)==24*13

# test <- data.frame(cbind(address, beds))
# test %>% filter(beds==1)
# rm(test)
rm(beds2, beds4, beds5, beds7, beds8, beds9, beds12)
```

## Number of Full Bath
```{r}
bathf <- c(scrape(coldwell_tc1, "li:nth-child(2) .val"), scrape(coldwell_tc2, "li:nth-child(2) .val"), scrape(coldwell_tc3, "li:nth-child(2) .val"), scrape(coldwell_tc4, "li:nth-child(2) .val"), scrape(coldwell_tc5, "li:nth-child(2) .val"), scrape(coldwell_tc6, "li:nth-child(2) .val"), scrape(coldwell_tc7, "li:nth-child(2) .val"), scrape(coldwell_tc8, "li:nth-child(2) .val"), scrape(coldwell_tc9, "li:nth-child(2) .val"), scrape(coldwell_tc10, "li:nth-child(2) .val"), scrape(coldwell_tc11, "li:nth-child(2) .val"), scrape(coldwell_tc12, "li:nth-child(2) .val"), scrape(coldwell_tc13, "li:nth-child(2) .val"))

bathf <- data.frame(bathf)
bathf <- bathf %>% mutate(bathf = as.numeric(bathf))
nrow(bathf)==24*13
```

## Number of Partial Bath
```{r}
bathp1 <- c(1, 2, 1, 1, 1, 0, 1, 1, 1, 0, 0, 1, 1, 0, 0, 0, 2, 1, 1, 2, 0, 1, 1, 0)
bathp2 <- c(0, 1, 0, 1, 1, 1, 0, 1, 0, 2, 0, 1, 0, 0, 1, 1, 1, 1, 1, 0, 1, 1, 1, 1)
bathp3 <- c(0, 1, 1, 3, 0, 1, 0, 1, 0, 2, 1, 0, 1, 0, 0, 1, 1, 1, 0, 1, 1, 0, 2, 0)
bathp4 <- c(2, 1, 0, 1, 2, 0, 2, 0, 1, 1, 0, 1, 0, 0, 0, 0, 0, 1, 0, 1, 1, 1, 1, 0)
bathp5 <- c(0, 3, 0, 1, 1, 0, 1, 0, 1, 2, 2, 1, 0, 0, 1, 1, 2, 3, 1, 1, 1, 1, 1, 1)
bathp6 <- c(2, 1, 0, 0, 0, 0, 1, 1, 1, 0, 0, 0, 0, 0, 1, 2, 0, 0, 1, 0, 1, 0, 0, 0)
bathp7 <- c(2, 1, 0, 0, 1, 0, 1, 1, 1, 1, 1, 0, 1, 1, 0, 0, 0, 1, 1, 0, 0, 1, 2, 1)
bathp8 <- c(1, 0, 0, 0, 0, 0, 2, 0, 0, 0, 0, 0, 3, 1, 0, 0, 0, 1, 0, 0, 0, 1, 2, 1)
bathp9 <- c(1, 0, 3, 0, 0, 0, 1, 2, 0, 1, 1, 0, 0, 1, 0, 0, 0, 1, 1, 0, 0, 0, 1, 0) 
bathp10 <- c(1, 0, 1, 0, 1, 1, 1, 1, 0, 1, 0, 1, 0, 0, 0, 1, 0, 1, 1, 1, 0, 1, 0, 0)
bathp11 <- c(2, 1, 0, 1, 5, 1, 1, 1, 1, 2, 0, 0, 1, 0, 1, 1, 1, 0, 0, 1, 2, 1, 0, 1)
bathp12 <- c(2, 2, 0, 2, 1, 1, 1, 1, 0, 0, 0, 1, 0, 0, 1, 1, 1, 0, 0, 0, 2, 0, 2, 0)
bathp13 <- c(0, 1, 3, 0, 1, 1, 0, 0, 0, 0, 0, 1, 0, 0, 2, 2, 1, 0, 0, 1, 1, 0, 1, 3)

# test <- data.frame(cbind(address, bathp))
# test
# rm(test)

bathp <- c(bathp1, bathp2, bathp3, bathp4, bathp5, bathp6, bathp7, bathp8, bathp9, bathp10, bathp11, bathp12, bathp13)
bathp <- data.frame(bathp)
nrow(bathp)==24*13
rm(bathp1, bathp2, bathp3, bathp4, bathp5, bathp6, bathp7, bathp8, bathp9, bathp10, bathp11, bathp12, bathp13)
```

## Number of Car Garage
```{r}
carg <- c(scrape(coldwell_tc1, ".car-garage .val"), scrape(coldwell_tc2, ".car-garage .val"), scrape(coldwell_tc3, ".car-garage .val"), scrape(coldwell_tc4, ".car-garage .val"), scrape(coldwell_tc5, ".car-garage .val"), scrape(coldwell_tc6, ".car-garage .val"), scrape(coldwell_tc7, ".car-garage .val"), scrape(coldwell_tc8, ".car-garage .val"), scrape(coldwell_tc9, ".car-garage .val"), scrape(coldwell_tc10, ".car-garage .val"), scrape(coldwell_tc11, ".car-garage .val"), scrape(coldwell_tc12, ".car-garage .val"), scrape(coldwell_tc13, ".car-garage .val"))

carg <- data.frame(carg)
nrow(carg)==24*13 
24*13-nrow(carg) # 29  

# length(scrape(coldwell_tc1, ".car-garage .val"))
# scrape(coldwell_tc6, ".street-address")

carg1 <- (scrape(coldwell_tc1, ".car-garage .val"))
carg1[18:24] <- c(0, carg1[18:23])

carg2 <- (scrape(coldwell_tc2, ".car-garage .val"))
carg2[20:23] <- c(0, carg2[20:22])
carg2[21:24] <- c(0, carg2[21:23])

carg3 <- (scrape(coldwell_tc3, ".car-garage .val"))
carg3[19:24] <- c(0, carg3[19:23])

carg4 <- (scrape(coldwell_tc4, ".car-garage .val"))
carg4[15:24] <- c(0, carg4[15:23])

carg5 <- (scrape(coldwell_tc5, ".car-garage .val"))
carg5[10:24] <- c(0, carg5[10:23]) 

carg6 <- (scrape(coldwell_tc6, ".car-garage .val"))
carg6[3:21] <- c(0, carg6[3:20]) 
carg6[10:22] <- c(0, carg6[10:21]) 
carg6[19:23] <- c(0, carg6[19:22]) 
carg6[23:24] <- c(0, carg6[23:23]) 

carg7 <- (scrape(coldwell_tc7, ".car-garage .val")) 
carg7[15:20] <- c(0, carg7[15:19]) 
carg7[17:21] <- c(0, carg7[17:20]) 
carg7[18:22] <- c(0, carg7[18:21]) 
carg7[20:23] <- c(0, carg7[20:22]) 
carg7[21:24] <- c(0, carg7[21:23]) 

carg8 <- (scrape(coldwell_tc8, ".car-garage .val")) 
carg8[1:24] <- c(0, carg8[1:23]) 

carg9 <- (scrape(coldwell_tc9, ".car-garage .val"))
carg9[5:21] <- c(0, carg9[5:20]) 
carg9[9:22] <- c(0, carg9[9:21]) 
carg9[12:23] <- c(0, carg9[12:22]) 
carg9[21:24] <- c(0, carg9[21:23]) 

carg10 <- (scrape(coldwell_tc10, ".car-garage .val"))
carg10[15:22] <- c(0, carg10[15:21])
carg10[21:23] <- c(0, carg10[21:22])
carg10[24:24] <- c(0)

carg11 <- (scrape(coldwell_tc11, ".car-garage .val"))
carg11[3:22] <- c(0, carg11[3:21])
carg11[5:23] <- c(0, carg11[5:22])
carg11[7:24] <- c(0, carg11[7:23])

carg12 <- (scrape(coldwell_tc12, ".car-garage .val"))
carg12[19:24] <- c(0, carg12[19:23])

carg13 <- (scrape(coldwell_tc13, ".car-garage .val"))
carg13[8:23] <- c(0, carg13[8:22])
carg13[9:24] <- c(0, carg13[9:23])

carg <- c(carg1, carg2, carg3, carg4, carg5, carg6, carg7, carg8, carg9, carg10, carg11, carg12, carg13)

carg <- data.frame(carg)
carg <- carg %>% mutate(carg = as.numeric(carg))
nrow(carg)==24*13

# test <- data.frame(cbind(address, carg))
# test %>% filter(carg==0)
# rm(test)

rm(carg1, carg2, carg3, carg4, carg5, carg6, carg7, carg8, carg9, carg10, carg11, carg12, carg13)
```

## Price
```{r}
price <- c(scrape(coldwell_tc1, ".price-normal"), scrape(coldwell_tc2, ".price-normal"), scrape(coldwell_tc3, ".price-normal"), scrape(coldwell_tc4, ".price-normal"), scrape(coldwell_tc5, ".price-normal"), scrape(coldwell_tc6, ".price-normal"), scrape(coldwell_tc7, ".price-normal"), scrape(coldwell_tc8, ".price-normal"), scrape(coldwell_tc9, ".price-normal"), scrape(coldwell_tc10, ".price-normal"), scrape(coldwell_tc11, ".price-normal"), scrape(coldwell_tc12, ".price-normal"), scrape(coldwell_tc13, ".price-normal"))

price <- data.frame(price)
price <- price %>%
  mutate(price = as.numeric(gsub("\\$|,", "", as.character(price))))
nrow(price)==24*13
```


# Data: Make the Data Table
```{r}
house <- cbind(address, cityzip, beds, bathf, bathp, carg, price)
# house <- house[-27, ] # remove duplicate
# save(house, file = "../Source/house.Rda")
# write_csv(house, "../Source/house.csv")
# house <- read_csv("../Source/house.csv")
```








# Getting Difficult Variables

# Make a List of lid
```{r message=FALSE, warning=FALSE}
## Load Data
## Temporary. Store in one eventually. Clean. !!!

house  <- read_csv("../Source/house.csv")
house2 <- read_csv("../Source/house2.csv")

house2 <- data.frame(house2) %>% select(address, sqft, pid)
house2$pid[45] <- "pid_21734880/"
house2 <- house2[-27,]

house3 <- house %>% left_join(house2, by = "address")

house3$pid[duplicated(house3$pid)]
duplicated(house3$pid)

house3$address[duplicated(house3$address)]
duplicated(house3$address)

# save(house3, file = "../Source/house3.Rda")
# write_csv(house3, "../Source/house3.csv")
# house3 <- read_csv("../Source/house3.csv")


## Make a List
lstlid <- house3 %>%
  mutate(first = paste(tolower(gsub("\\s", "-", city)), "/"),
         second = paste(tolower(gsub("\\s", "-", address)), "/"),
         lid = gsub("\\s", "", 
                    paste("https://www.coldwellbankerhomes.com/mn/", first, second, pid))) %>%
  select(lid)
```


# Inventory
```{r}
lidvec <- lstlid$lid
soldat <- "ul:nth-child(4) li:nth-child(2)"
listedat <- "body > section.content > div:nth-child(1) > div.layout-main.property-details > div:nth-child(8) > div > div > div.col-1 > ul:nth-child(4)"
schname <- ".school-name"
schaddr <- ".school-address"
schgd <- ".property-detail-school-listing .school-data-grades"
schdist <- ".property-detail-school-listing .school-data-distance"
schdstr <- ".split-2 .col:nth-child(2) li" #?
```


# Function to create a vector of a variable from a given path
```{r}
scrape2 <- function(path) {
  megalst <- c()
  for (i in 1:length(lidvec)) {
    upd <- scrape(read_html(lidvec[i]), path)
    megalst <- c(megalst, upd)
  }
  return(megalst)
}
```

# Function to Saefely Scrape
```{r}
safely <- function(fn, ..., max_attempts = 10) {
  function(...) {
    this_env <- environment()
    for(i in seq_len(max_attempts)) {
      ok <- tryCatch({
          assign("result", fn(...), envir = this_env)
          TRUE
        },
        error = function(e) {
          FALSE
        }
      )
      if(ok) {
        return(this_env$result)
      }
    }
    msg <- sprintf(
      "%s failed after %d tries; returning NULL.",
      deparse(match.call()),
      max_attempts
    )
    warning(msg)
    NULL
  }
}
```

# New Function
```{r}
safe_scrape2 <- safely(scrape2)
```


# More Variables

## Sold at
```{r}
# soldatvar <- safe_scrape2(soldat)

soldatpr_date <- soldatvar %>%
  mutate(soldatpr = as.numeric(gsub("[^0-9\\.]", "", substr(soldatvar, 10, 19))),
         soldatdate = as.Date(gsub("/18", "/2018", 
                                   paste0("0", gsub("\\s", "", 
                                                    sub('.*\\on', '', soldatvar)))), "%m/%d/%Y")) %>%
  select(soldatpr, soldatdate)
soldatpr_date <- soldatpr_date[-27, ]

# save(soldatpr_date,  file = "../Source/soldatpr_date.Rda")
```

## Listed at
```{r}
listedatvar <- safe_scrape2(listedat)
listedatvar <- data.frame(listedatvar)
nrow(listedatvar) # only 221 houses available

tem <- listedatvar %>% # listedat price and date for only 221 houses
  mutate(var = gsub("\r\n", "", sub('.*\\Listed at', '', as.character(listedatvar) ))) %>%
  select(var)

listedatpr <- listedatvar %>%
  mutate(var = gsub("\r\n", "", sub('.*\\Listed at', '', as.character(listedatvar) )),
         listedatpr = as.numeric(gsub("[^0-9\\.]", "", substr(var, 2, 11)))) %>%
  select(listedatpr)


## Note: Do not have listed at date variable yet due to complicated string conversion


## Necessary adjustment due to the absence of listedat price
### 74. 1718 Highland Parkway (no listed at) only Sold at $415,000 on 1/23/18
listedatpr$listedatpr[74] <- 415000

## Remove duplicate
listedatpr <- soldatpr[-27, ]


## Notes
gsub("[^0-9\\.]", "", substr(a$var, 2, 11))
a <- sub('.*\\on', '', a$var)
b <- gsub("\\s", "", a)


## Still working on it
ifelse(b[length(a)-1:length(a)]=="18", gsub("/18","/2018", substr(4, nchar(a), a)), gsub("/17","/2017", substr(4, nchar(a), a)))
# as.Date(gsub("/18", "/2018", paste0("0", gsub("\\s", "", sub('.*\\on', '', var)))), "%m/%d/%Y"))
```

## School Name
```{r}
schnamevar <- safe_scrape2(schname)
# save(schnamevar,  file = "../Source/schnamevar.Rda")
```

## School Address
```{r}
schaddrvar <- safe_scrape2(schaddr)
# save(schaddrvar,  file = "../Source/schaddrvar.Rda")
```

## School Grade
```{r}
schgdrvar <- safe_scrape2(schgd)
# save(schgdrvar,  file = "../Source/schgdrvar.Rda")
```

## School Distance
```{r}
schdistvar <- safe_scrape2(schdist)
```

## School District
```{r}
schdstrvar <- safe_scrape2(schdstr)
```



# Load and Combine
```{r message=FALSE, warning=FALSE}
load("../Source/soldatpr_date.Rda")
house3 <- cbind(house3, soldatpr_date)
# save(house3, file = "../Source/house3.Rda")
# write_csv(house3, "../Source/house3.csv")
# house3 <- read_csv("../Source/house3.csv")
```



# Unused
### GOLDEN! ####
```{r}
# Scraping pid
# scrape_pid <- function(readhtml, path) {
#   varlst <- readhtml %>%
#     html_nodes(path) %>%
#     html_attr("data-pid") 
#   return(varlst) }

# Notes
## . for class
## # for id

# scrape_pid(coldwell_tc1, ".property-snapshot-psr-panel")
```