---
title: "| Survival Analysis    \n| Data Analysis Version 1\n"
author: "April Leclair"
date: "2018-02-25"
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "Data Analysis - HTML Output") })
output:
  bookdown::tufte_html2:
    number_sections: no
    split_by: none
    toc: no
  bookdown::pdf_document2:
    latex_engine: pdflatex
    number_sections: no
    toc: no
  bookdown::tufte_handout2:
    latex_engine: xelatex
    number_sections: no
    toc: no
  bookdown::html_document2:
    number_sections: no
    split_by: none
    toc: no
---
 
```{r setup, include=FALSE, message=FALSE}
library(tidyverse)
library(readr)
library(ggplot2)
library(survival)
library(survminer)
knitr::opts_chunk$set(tidy = FALSE, message=FALSE, cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)
```


### Load Data
```{r message=FALSE, warning=FALSE}
house <- read_csv("../Source/house_TC.csv")

```


### Variable Names
```{r}
names(house)
```


### Treat Time = 0
```{r}
# Note: Only 1 case has `tom` = 0
house <- house %>% mutate(tom = ifelse(tom == 0, 1, tom))
```


### Graphs: Exponential - density, CDF, and survival curve
```{r}
survreg( Surv(tom) ~ 1 , dist = "exponential" , data = house )
```

```{r}
# Density
house %>% ggplot(aes(x=tom)) + 
  stat_function( fun=dexp, args= 1/exp(4.454011)) + 
  labs(title = "Exponential PDF for Time on Market", x="Time on Market", y="Density",
       caption = "Data collected from Coldwell Banker")

# CDF
house %>% ggplot(aes(x=tom)) + 
  stat_function(fun=pexp, args= 1/exp(4.454011)) + 
  labs(title = "Exponential CDF for Time on Market", x="Time on Market", y="Cumulative Proportion",
       caption = "Data collected from Coldwell Banker")

# Survival
house %>% ggplot(aes(x=tom)) + 
  stat_function(fun=pexp, args= 1/exp(4.454011), lower.tail=FALSE) + 
  labs(title = "Exponential Survival Curve for Time on Market", x="Time on Market", y="Survival Proportion",
       caption = "Data collected from Coldwell Banker")
```


### Graphs: Weibull - density, CDF, and survival curve
```{r}
survreg( Surv(tom) ~ 1 , dist = "weibull" , data = house )
```

```{r}
# Density
house %>% ggplot(aes(x=tom)) + 
  stat_function( fun=dweibull, args=list(shape=1/0.5511901, scale=exp(4.575712))) + 
  labs(title = "Weibull PDF for Time on Market", x="Time on Market", y="Density",
       caption = "Data collected from Coldwell Banker")

# CDF
house %>% ggplot(aes(x=tom)) + 
  stat_function(fun=pweibull, args=list(shape=1/0.5511901, scale=exp(4.575712))) + 
  labs(title = "Weibull CDF for Time on Market", x="Time on Market", y="Cumulative Proportion",
       caption = "Data collected from Coldwell Banker")

# Survival
house %>% ggplot(aes(x=tom)) + 
  stat_function(fun=pweibull, args=list(shape=1/0.5511901, scale=exp(4.575712)), lower.tail=FALSE) + 
  labs(title = "Weibull Survival Curve for Time on Market", x="Time on Market", y="Survival Proportion",
       caption = "Data collected from Coldwell Banker")
```


### Chance that the house will be on the market for more than 50 days
```{r}
# Exponential
1-pexp(50, 1/exp(4.454011))

# Weibull
1-pweibull(50, shape=1/0.5511901, scale=exp(4.575712))
```


### Mean time to sell
```{r}
# Exponential
S = function(x) 1-pexp(x, 1/exp(4.454011))
integrate(S,0,Inf)

# Weibull
S = function(x) 1-pweibull(x, shape=1/0.5511901, scale=exp(4.575712))
integrate(S,0,Inf)
```


### Median time to sell
```{r}
qexp(0.5, 1/exp(4.454011))
qweibull(0.5, shape=1/0.5511901, scale=exp(4.575712))
```


## Kaplan-Meier curves

### Kaplan-Meier curves with CI - all groups
```{r}
KM = survfit( Surv(tom) ~ 1 , conf.type="plain" , data=house ) 
ggsurvplot(KM, data=house) +
  labs(x = "Days on Market", y = "Survival Probability", title = "Kaplan-Meier Curve of Days on Market",
       caption = "Data collected from Coldwell Banker")
```


### Kaplan-Meier curves with CI - by price difference groups
```{r}
# Note: Price Difference = sold - listed
KMECDF = survfit(Surv(tom) ~ prdiff_np, data=house )
ggsurvplot(KMECDF, data=house, legend.title="Type", 
           legend.labs=c("Positive Price Difference","Negative Price Difference", "No Price Difference")) +
    labs(x = "Days on Market", y = "Survival Probability", title = "Kaplan-Meier Curve of Days on Market",
       caption = "Data collected from Coldwell Banker")
```

### Inspect Confidence Interval for Each Group
```{r}
summary(KMECDF)
```


### KM Median 
```{r}
# median for all groups
KM

# median by price difference groups
KMECDF
```

### KM Mean Survival
```{r}
# Finds area under Kaplan-Meier curve (if largest observation is censored, this assumes that the K-M drops to 0 at that value).

# AUCKM stands for "Area Under Curve Kaplan Meier":
AUCKM = function(survobj,duration)
{
base=c(0,summary(survobj)$time,max(duration))
heights=c(1,summary(survobj)$surv)
new=c()
for(i in 1:length(heights)) { new=c(new,(base[i+1]-base[i])*heights[i]) }
c(sum(new))
}

# mean survival for all groups
AUCKM(KM, house$tom)

# mean survival by price difference groups
AUCKM(KMECDF, house$tom)
```


## Actuarial Method Estimate

### Life Table Function 
```{r}
LifeTable = function( time, status, breaks, finite ) {
  failed = c(0,hist(time[status==1],breaks=breaks,plot=F)$counts)
  censored = c(0,hist(time[status==0],breaks=breaks,plot=F)$counts)
  alivestart = length(time)-cumsum(failed[-length(failed)])-cumsum(censored[-length(censored)])
  atrisk = alivestart-c(censored[-1]/2)
  failrate = failed[-1]/atrisk
  survrate = 1-failrate
  survest = c(1,cumprod(1-failrate)[1:(length(failrate))])
  if (finite == 0) 
    return(as.data.frame(cbind(Failed = failed[-1], Censored=censored[-1], AliveStart=alivestart,
                             AtRisk = atrisk[-length(atrisk)], FailRate = failrate[-length(failrate)],
                SurvRate = survrate[-length(survrate)], SurvEst = survest[-length(survest)])))
  if (finite == 1) 
    return(as.data.frame(cbind(Failed = failed[-1], Censored = censored[-1], AliveStart = alivestart, 
                               AtRisk = atrisk, FailRate = failrate, SurvRate = survrate, SurvEst = survest))) }
```

### Plot Survival Estimates from Life Table
```{r}
status <- rep(1, 311)
b = c( 0 , 25 , 50 , 75 , 100 , 125 , 150 , 175 , 200 , 225 , 250 , 300 , 325 , 350, 375)
(l = LifeTable(house$tom, status, breaks=b, finite=1))
# !!! number of rows of result is not a multiple of vector length (arg 1)

step = stepfun( c(25,50,75,100,125,150,175,200,225 , 250 , 300 , 325 , 350, 375) , l$SurvEst )
plot( step , do.points=FALSE , ylab="Survival" , xlab="Months" , main="")
```

#### Plot KM & Actuarial method estimate for Comparison
```{r}
# ggplot!!! both plots
# KM = survfit( Surv(tom) ~ 1 , conf.type="plain" , data=house ) 
# ggsurvplot(KM, data=house) +
#   labs(x = "Days on Market", y = "Survival Probability", title = "Kaplan-Meier Curve of Days on Market",
#        caption = "Data collected from Coldwell Banker")

plot(KM, ylab="Survival" , xlab="Days" , main="House Survival")
plot( step , do.points=FALSE, add=TRUE, col = "red")
```

### Actuarial estimate of the hazard function
```{r}
HazardEst = l$FailRate/25
hazstep = stepfun( c(0,25,50,75,100,125,150,175,200,225,250,300,325,350,375,400) , c(0,HazardEst,0) )
plot( hazstep, do.points=FALSE, ylab="h(k)", xlab="k", main="", xlim=c(0,225) )
```



### Plot Weibull: Compare estimated hazard and survival functions to those obtained via the actuarial method
```{r}
(weib <- survreg( Surv(tom) ~ 1 , dist='weibull' , data=house))
```

#### We can add the estimated Weibull hazard to the graph of the piecewise constant actuarial hazard estimate:

```{r}
plot(hazstep,do.points=FALSE,ylab='h(k)',xlab='k',main="",xlim=c(0,225))
curve(dweibull(x, shape=1/0.5511901, scale=exp(4.575712)) / (1-pweibull(x, shape=1/0.5511901, scale=exp(4.575712))), add=T, col='red')
```

```{r}
g = ggplot(data=data.frame(x=c(-1, 0,25,50,75,100,125,150,175,200,225,250,300,325,350,375,400),
                       y=c(0,HazardEst,0)) , aes(x=x,y=y)) +
  geom_step() +
  labs(x='k',y='h(k)')

hweibull = function(x,shape,scale) { dweibull(x,shape=shape,scale=scale)/(1-pweibull(x,shape=shape,scale=scale)) }

g + stat_function(fun=hweibull,args=list(shape=1/0.5511901, scale=exp(4.575712)) , col='red') 
```


## Hazard Functions

### Exponential
```{r}
hexp = function(x,lambda) { dexp(x,lambda) / ( 1-pexp(x,lambda) ) }
```

```{r}
curve( hexp( x , 1/exp(4.454011) ) , from=0 , to=max(house$tom) , ylim = c(0,0.02), col = "red")
```


### Weibull
```{r}
hweibull = function(x, a, b) { dweibull(x, a, b) / ( 1-pweibull(x, a, b) ) }
```

```{r}
curve( hweibull( x , 1/0.5511901, exp(4.575712)) , from=0 , to=max(house$tom) , ylim = c(0,0.06), col = "red")
```


