---
title: "Midterm Paper"
author: "Nick McMullen, April Choi"
date: "March 5, 2018"
output:
  pdf_document:
    pandoc_args: [
      "-V", "classoption=twocolumn"
    ]
  html_document: default
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readr)
library(ggplot2)
library(survival)
library(survminer)

house <- read_csv("../Source/house_TC.csv")

house <- house %>% mutate(tom = ifelse(tom == 0, 1, tom))

house$prdiff_np <- as.factor(house$prdiff_np)

```




## 1 | Introduction


### 1.1 Background

The real estate market in the Twin Cities (Minneapolis and St. Paul, MN, USA), and many places across the United States, has been thriving on the economic recovery since the end of the Great Recession. While the current market is, to the casual observer, favorable to sellers some houses in Twin Cities still take a surprisingly long time to sell. We will investigate the characteristics that impact time on market (TOM) for houses in the Twin Cities through a survival analysis, or time-to-event, framework.


### 1.2 Review of Literature

\textbf{Price}

One of the most obvious factors that may influence TOM for any house is price. A high price may scare potential buyers away, while a low price might invite skepticism or a long bidding war. The "right" pricing decision is a difficult one to make for sellers and agents, but reaching an equilibrium between buyer and seller satisfaction may be the best way to minimize TOM. Cheng, Lin, and Liu (2008) developed a closed-form formula to uncover the theoretical relationship between price and TOM. They aim to describe the marginal benefit of keeping a house on the market longer.

The authors used house sales data from Fannie Mae and Freddie Mac, U.S. federal mortgage agencies, to identify a nonlinear positive relationship between price and TOM. They approached this analysis by assuming that a buyer and seller arrive at an agreed upon price following the Poisson processes at rate $\lambda$. While utilizing methods from Bond et al. (2007), they investiaged some assumptions about the shape of TOM data, including normal, chi-square, Weibull, and exponential distributions. Like Bond et al. (2007), Cheng, Lin, and Liu (2008) found that the exponential distribution best fit the TOM data. 

The findings from Cheng, Lin, and Liu's (2008) investigation was that marginal benefit on sale price decreases after each offer that is made on a house. As a house's price approaches equilibrium between seller and buyer satisfaction, it does the seller little good to leave the house on the market longer.

\textbf{Duration Dependence}

Another interesting question regarding TOM is how the probability of sale changes with time. Thomas W. Zuehlke (1987) conducted research on this very topic. He used 290 single family detached homes obtained from a 1982 multiple listing service (MLS) book in Tallahassee, FL, USA and observed how likely each house was to sell based on its status of vacant or occupied. Zuehlke utilized a Weibull hazard model to compare the two and found that stronger incentives existed for sellers of vacant houses to reduce prices faster than those of occupied homes. Thus, Zuehlke finds that vacant houses tend to exhibit positive duration dependence, while occupied homes showed little evidence of duration dependence. 

Motivated by a Massachusetts state policy adopted in 2006 that prohibits sellers from resetting their home's TOM by relisting, Tucker, Zhang, and Zhu (2013) also investigated the impact of TOM on sale price. The authors agree that longer TOM is negatively associated with buyer perception, which lines up with the Massachusetts relisting policy. To investigate if TOM was indeed a significant deterrent for buyers looking at an otherwise appealing house, the authors analyzed the TOM for homes before and after the enactment of the policy. They obtained listings data for residential properties on the market between January 2005 and June 2007 from two MLSs: MLS-PIN, which serves Massachusetts, and the State-Wide Rhode Island MLS. Their analysis consisted of three groupings of homes: listed and sold before the policy change, listed before and sold after, and listed and sold after. The authors utilized simple linear regression between the three groups and found that the homes listed before the policy and sold after were most severely impacted, resulting in an average sale price reduction of $16,000. 

\textbf{Listing Price Changes}

Listing price changes may also affect the TOM for a house. A listing price change could attract more attention to a house on the market, but could also indicate a longer TOM. John R. Knight (2002) conducted an analysis on listing price changes, investigated which types of homes were most likely to go through multiple listing price changes, and which price changes tend to give the worst results for the seller.

Knight used 3490 detached single family homes that sold between January 1997 and December 1998 by Metroservices Inc. of Sacramento, CA, USA. He notes that price adjustment data is generally missing from TOM datasets, but his study incorporated this into the analysis to examine the determinants of list price changes. He utilized a maximum likelihood probit model and found that the two most important determinants of price changes are TOM and initial markup. Atypical homes tend to not see significant price changes because there is little market precedent for price changes after a certain TOM. Knight's findings are consistent with previous research on pricing under demand uncertainty. 


### 1.3 Research Question (and mapping paragraph)

We are interested in which factors affect the TOM for a house in the Twin Cities. Some variables of interest include the amount of recorded crime in a house's designated neighborhood and the proximity of a house to the nearest school. 

In the following sections, we outline our data collection and methods and provide a brief descriptions of our assumptions. 

##2 | Data and Methods

### 2.1 Data and Sources

#### Real Estate Data

Our data represents a random sample of 311 single family homes in the Twin Cities. A search of all recently sold homes in the area was conducted in the Coldwell Banker MLS and 311 unique addresses were randomly selected for the project. Some variables were scraped directly from the Coldwell Banker site, while others were manually inputted. 

#### Crime Data

We also obtained crime data beginning on January 1, 2015 from the government websites for the cities of Minneapolis and St. Paul. Each crime was matched with a neighborhood, aggregated into total neighborhood crime counts, and then joined with the house sales data. This addition will provide insight on the number of crimes that have been recorded over the past three years near each sold house. 

### 2.2 Variables

For all 311 observations, we collected unique addresses `address`, city of residence `city`, zip code `zip`, number of bedrooms `beds`, number of full bathrooms `bathf`, number of partial bathrooms `bathp`, number car garage `carg`, house squarefeet `sqft`, price the home was listed at `listedatpr`, price the home was sold for `soldatpr`, date the home was listed `listedatdate`, date the home was sold `soldatdate`, neighborhood the house is located `neighb`, distance to nearest school (mi.) `schdist`, and number of crimes that were recorded in that neighborhood beginning on January 1, 2015 `ncrime`.

### 2.3 Assumptions

An essential assumption to our research is based on censoring. We intentionally collected our data from a sample of recently sold houses to limit censoring. If we were to take a true random sample of all houses listed for sale in the past $X$ number of years, we realize that a significant amount of right censoring could be present, because at the time of data collection some houses in our sample would have not been sold yet. We acknowledge that sampling from a set of recently sold houses may introduce some bias into our results - all recently sold houses (late January to early Februray 2018) may share certain unknown or unaccounted-for characteristics that made them sell. This bias is acceptable considering the convenience and added accuracy provided by uncensored data.

### 2.4 Methods

We utilize visualizations of parametric and non-parametric estimates of the true survival curve for `Time on Market (tom)` and test for significant differences in `tom` between houses with different characteristics, such as difference between list and sold price and neighborhood, using the Mantel-Cox test statistic in a log-rank test. We also test for significant relationships between `tom` and other quantitative variables such as `Number of neighborhood crimes in last 2 years (ncrime)`, `Distance from the nearest school (schdist)`, `list price (listedatpr)`, and `sale price (soldatpr)`.

##3 | Preliminary Results

We begin our preliminary analysis by investigating the length of time that houses in the Twin Cities remain on the market after their preliminary listing date.



### 3.1 Kaplan-Meier Curves

#### Aggregate

First, a Kaplan-Meier curve with confidence intervals for survival (time on market) across all groups.

```{r echo=FALSE}

KM = survfit( Surv(tom) ~ 1 , conf.type="plain" , data=house ) 
ggsurvplot(KM, data=house) +
  labs(x = "Days on Market", y = "Survival Probability", title = "Kaplan-Meier Curve of Days on Market")

```

#### Categorical Price Differences 


Next, a Kaplan-Meier curve stratified for three groupings of sold houses: houses that sold below list value, at list value, and above list value.

```{r echo=FALSE}

# Note: Price Difference = sold - listed
KMECDF = survfit(Surv(tom) ~ prdiff_np, data=house )
ggsurvplot(KMECDF, data=house, legend.title="Type", 
           legend.labs=c("Positive Price Difference","Negative Price Difference", "No Price Difference")) +
    labs(x = "Days on Market", y = "Survival Probability", title = "Kaplan-Meier Curve of Days on Market - Stratified")

```

We see that houses in our data that sold for under the listed price tended to take the longest to sell, while houses that sold for above the listed price tended to sell more quickly. This result is intuitive, as houses that are priced competitively tend to sell faster while houses that see price reductions are likely to have been on the market longer. 


### 3.2 Parametric Survival Curve Estimates

#### Exponential Aggregate

A survival curve for all houses in the data estimated parametrically by the exponential distribution.

```{r echo=FALSE, include=FALSE, message=FALSE}

survreg(Surv(tom) ~ 1, dist = "exponential", data=house)

```


```{r echo=FALSE, message=FALSE, warning=FALSE}

surv <- function(x){1-pexp(x, 1/exp(4.454011))}

house %>% ggplot(aes(x=tom)) + 
  stat_function(fun=surv, lower.tail=FALSE) + 
  labs(title = "Exponential Survival Curve for Time on Market", x="Time on Market", y="Survival Proportion",
       caption = "Data collected from Coldwell Banker")

```


####Exponential - Categorical Price Differences

```{r echo=FALSE, include=FALSE, message=FALSE}

survreg(Surv(tom) ~ as.factor(prdiff_np), dist = "exponential", data=house)

```


```{r echo=FALSE, message=FALSE, warning=FALSE}

surv <- function(x){1-pexp(x, 1/exp(4.0672313))}
surv1 <- function(x){1-pexp(x, 1/exp(4.0672313 + 0.5522913))}
surv2 <- function(x){1-pexp(x, 1/exp(4.0672313 + 0.1166739))}


house %>% ggplot(aes(x=tom)) + 
  stat_function(fun=surv, lower.tail=FALSE, col="black") + 
  stat_function(fun=surv1, lower.tail=FALSE, col="red") + 
  stat_function(fun=surv2, lower.tail=FALSE, col="blue") +
  labs(title = "Exponential Survival Curve for Time on Market Stratified for Price Difference", x="Time on Market", y="Survival Proportion",
       caption = "Data collected from Coldwell Banker")


```



#### Weibull - Aggregate

```{r echo=FALSE, include=FALSE, message=FALSE}

survreg(Surv(tom) ~ 1, dist = "weibull", data=house)

```


```{r echo=FALSE, message=FALSE, warning=FALSE}

surv <- function(x){1-pweibull(x, shape=1/.5511901, scale=exp(4.575712))}

house %>% ggplot(aes(x=tom)) + 
  stat_function(fun=surv, lower.tail=FALSE) + 
  labs(title = "Weibull Survival Curve for Time on Market", x="Time on Market", y="Survival Proportion",
       caption = "Data collected from Coldwell Banker")

```


####Weibull - Categorical Price Differences

```{r echo=FALSE, include=FALSE, message=FALSE}

survreg(Surv(tom) ~ prdiff_np, dist = "weibull", data=house)

```

```{r echo=FALSE, message=FALSE, warning=FALSE}

surv3 <- function(x){1-pweibull(x, shape=1/0.5340263, scale=exp(4.2485166))}
surv4 <- function(x){1-pweibull(x, shape=1/0.5340263, scale=exp(4.2485166 + 0.4589256))}
surv5 <- function(x){1-pweibull(x, shape=1/0.5340263, scale=exp(4.2485166 + 0.1177328))}


house %>% ggplot(aes(x=tom)) + 
  stat_function(fun=surv3, lower.tail=FALSE, col="black") + 
  stat_function(fun=surv4, lower.tail=FALSE, col="red") +
    stat_function(fun=surv5, lower.tail=FALSE, col="blue") +  labs(title = "Weibull Survival Curve for Time on Market", x="Time on Market", y="Survival Proportion",
       caption = "Data collected from Coldwell Banker") 
```

We observe again from this parameterization that houses that sold for more than the original list price tend to sell the fastest, while houses that underwent price reductions tend to remain on the market the longest. 

We have visualized non-parametric and parametric estimates of the true survival curve but is there truly a difference between groupings of houses with different categorical characteristics? 

### 3.3 Tests for Statistical Significance

#### List and Sale Price Difference

In order to determine if the differences in the estimated survival curves for this category are statistically significant, we conduct a log-rank test. The results are as follows:

$H_0: S_1(t) = S_2(t) = S_3(t) \ \forall \ t$

$H_A: S_1(t) \neq S_2(t) \neq S_3(t) \ \text{for some} \ t$


```{r echo=FALSE, include=FALSE}

survdiff(Surv(tom) ~ prdiff_np, data=house)

```

A log-rank test reveals a Mantel-Cox test statistic of $53.87$. This produces a p-value of nearly 0:

```{r echo=FALSE}

1 - pchisq(53.87, df=2)

```

Thus, we confidently reject the null hypothesis and conclude that the three survival curves for price difference (negative, zero, positive) differ for some `tom`. 







